\section{Introduction}\label{sec:intro}
Assessment is one of the most fundamental aspects of education, playing an integral role in teaching and learning. In particular, summative assessment is used for gauging to which extent learning objectives are being met. Adequately determining a student's proficiency of concepts, skills and competencies is a complex undertaking. Valid, objective, reliable, and fair assessments call for a setting in which students are given ample opportunity to provide evidence of their knowledge and skills, and that a teacher can effectively observe and measure their level of mastery and attainment of learning objectives\todo{attainment level reached}.

Computer Science is a multifaceted subject, bringing together conceptual (CS) knowledge, computational thinking, and problem-solving skills. This knowledge can be categorized as factual knowledge (knowing what), procedural knowledge (knowing how), conceptual knowledge (knowing why), and meta-cognitive knowledge (knowing about knowing)\cite{streun2001kennis}. Teachers need to assess these achievement levels on all these aspects. This is far from trivial. In addition to teachers' pedagogical and content knowledge, it requires knowledge on constructing frameworks, test specifications, and the ability to align learning objectives with assessment tasks which match specific classroom needs. This research explores the characteristic features of effective CS assessment in the domains of Algorithms and Physical Computing, which incorporate the wide range of knowledge, computational thinking and problem solving skills, and what pedagogical-content-knowledge teachers need for its implementation.




\subsection{Curricular Reforms}


As of 2019, a reformed Computing Science (CS) curriculum will become effective in the Netherlands. The reformed program follows in the footsteps of recent international reforms (such as the U.S. CSTA, U.K. National curriculum, Israel, New Zealand's new "Digital Technologies") where a shift from computer and ICT applications towards rigorous computing has taken place. Internationally, multiple reports (\citeA{Brown2013} in the UK, \citeA{Bell2014} in New Zealand, \citeA{Yadav2016} in the U.S.) indicate that teachers face multiple challenges whilst implementing new standards. Specifically, teachers struggle with assessment \citeA{Yadav2016}. In their research, \citeA{GroverPea2013} describe that paying attention to assessment is a pre-requisite to successful implementation of any curriculum and conclude that large deficiencies exist pertaining to assessment. \todo{DEZE PARAGRAAF FIXEN}

\subsection{Current Situation in the Netherlands}
In the Netherlands, the reform has resulted in a stronger focus on fundamental concepts, including analysis of standard algorithms and being able to compare efficiency and implementability \cite{Barendsen2016}. Higher-order (Computational Thinking) skills are positioned as a fundamental pillar, characterized as one of the 21st-century skills competencies \cite{SLO2015}. The idea of Computational Thinking (CT) presents a different way of thinking about computing proficiency by emphasizing the (problem-solving) skills that students need to solve complex real-world problems and evaluate proposed solutions. It also implies that solely measuring CS proficiency as the acquisition of core content knowledge no longer suffices \cite{Yadav2015}.




Currently, only a small percentage of the Dutch CS teachers has completed a full academic qualification to teach the subject \cite{tolboom2014informatica}.
All-in-all, little research has been done on the effectiveness of the assessment materials available \cite{Yadav2016}. Yet the new situation calls for teachers well-equipped PCK in order to cope with the changing requirements.\todo{ARTICLE FRIJDAG MIDDAG HIER ZIN UIT PLAKKEN}
%How will this fundamental shift evolve (in) the Dutch educational landscape?\todo{Erik: kopje? wat bedoel je hiermee Erik?}



New curricular material comes with new misconceptions. \citeauthor{duncan2017teachers}'s results confirm their concern and warn that insufficiently prepared teachers may not pick up student misconceptions, or may even introduce incorrect ideas themselves. Speculating on their content knowledge (as a component of the PCK), Dutch teachers currently focus primarily on designing and implementing computer systems and programs \cite{Schmidt2007}, rather the more fundamental underlying concepts which the new curriculum now focusses on. In the Netherlands there is no centralized or standardized exam for the subject, no validated assessment tasks, and no way to ascertain the effectiveness of teaching. Teachers must develop their own assessments. From their community of practice observations, \citeA{hermansen2014reworking} conclude that teachers create their own assessments, generally by adapting others. There is usually only one teacher in a school, leaving them isolated from their peers. Compared to initiatives such as the U.K. Computing At School (CAS), in the Netherlands professional development opportunities, communities of practice, sharing of experiences, resources, and lessons learned are sparse. All in all, teachers have no assessment benchmark, and lack valid and reliable example assessments \cite{Yadav2015}, and furthermore lack evaluation or reflection on their teaching strategies via peer-collaboration.




\subsection{Assessment}
Despite the proliferation of CT-focused programs throughout several curricula around the world, there is still no general method\todo{Erik: is die er wel in andere vakken?} for the easy development of effective CT assessments \cite{catete2017framework}. The effectiveness of an assessments encompasses, among other criteria, validity, transparency, fairness, fit-for-purpose and cost-effectiveness.  In their extensive review study, \citeA{LyeKoh2014} found a few studies which examine computational skills as outcomes. No studies were found focussing on upper secondary high school. We were able to locate one: \citeA{Lee2011}.

\citeA{denning2017remaining} notes that educators internationally have many basic questions about teaching and assessing CT and expresses his concerns for ineffective teaching. Along the same line, \citeA{2013Seiter} describe assessment models incorporating progression pathways as vital for establishing research-based computational curricula.


There is a gap between the (meager) research and classroom practice \cite{Yadav2015}. Computing education and research suffer from the lack of assessment instruments \cite{voogt2017effecten}. In their research \citeA{2010TewGuzdial}, argument the importance of validated assessment that could be widely applicable across tertiary curricular approaches. This need may be even more true for secondary education, where far less relevant research has been done in the first place. Also, research developments are not accessible to teachers (either they need a paid subscription, academic literature may be written incomprehensibly, or not practical as it can't be implemented in the classroom in a cost-effective manner). What is needed are validated assessments which are of high quality, sound, scientifically grounded, yet which can be implemented in a practical manner in a classroom setting.

With all these factors in place, what is needed for teachers and their PCK to adapt to, and successfully implement, the curriculum reform?

\subsection{Main objectives of this thesis}


Although many secondary education curricula have reformed their Computer Science objectives, teachers lack the PCK to assess specific topics such as Algorithms and Physical Computing. The transition from 'using' to 'creating' digital artifacts has its impact on learning objectives. On one hand the shift has led to the prevalence of more fundamental and theoretical computational concepts, on the other hand the need for Computational Thinking skills such as computational practices has become imminent for the analysis and design of digital artifacts. These CT skills emphasize processes over facts, and open-ended solutions rather than exact answers, making the development of effective assessments strategy far from trivial \cite{rodriguez2017assessingCTunplugged}.

Via educational design research, this thesis pursues to clarify how to assess (the design of) Algorithms and Physical Computing from a secondary education Computer Science perspective. We identify the relevant fundamental computational concepts and computational practices, and investigate and evaluate how attainment levels can be identified and measured through assessment. This results in student competency models for Algorithms and Physical Computing. Accordingly, the implications are explored from the perspective of computer science teachers. In order to prepare teachers to implement and deliver the assessments, we analyze and typify the characteristics of PCK for Physical Computing and Algorithms.


\subsection{Structure of this document}
This document is organized as follows
In section \ref{sec:background} we summarize the results of our literature research: what is known about (the assessment of) Computational Thinking, Algorithms and Physical Computing. In section \ref{sec:aim} we formulate our research questions. Section \ref{sec:significance} describes the educational and scientific significance of this research. Subsequently, section \ref{sec:method} describes the approach to answering those research questions. 